Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}

### 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Update to the manuscript version to be submitted

\subsubsection{Phospholipid bilayer with low hydration level}
Fig.~\ref{ordPhydr} shows the published~\cite{dvinskikh05b,ulrich94,bechinger91} experimental order parameters
for the glycerol backbone and choline as a function of hydration level.
Despite slight differences in temperature and acyl chain composition,
the three independently reported data sets for the choline ($\beta$ and $\alpha$) segments agree well with each other:
Both order parameters increase with decreasing hydration level.
The glycerol backbone order parameters (g$_{3}$, g$_{2}$, g$_{1}$), in contrast,
have been observed~\cite{dvinskikh05b} to slightly decrease with dehydration.
Note that the original experiments~\cite{dvinskikh05b,ulrich94,bechinger91} measured only 
absolute values, but here
we included the signs measured in separate studies~\cite{hong95a,hong95b,gross97}. 
Consequently, the negative $\beta$ order parameter actually increases with dehydration 
as its absolute value decreases~\cite{dvinskikh05b,ulrich94,bechinger91}.
\begin{figure}[]
  \centering
  \includegraphics[width=7.3cm]{../DATAreportediINblog/dehydration_lp.pdf}
%    \includegraphics[width=7.3cm]{../DATAreportediINblog/dehydration_smooth.pdf}
    \caption{\label{ordPhydr}
    The effect of dehydration on glycerol and choline order parameters in experiments.
    The magnitudes of order parameters are measured for DMPC ($^1$H-$^{13}$C NMR) at 314~K~\cite{dvinskikh05b}, 
    for POPC ($^2$H NMR) at 296~K~\cite{bechinger91} and for DOPC ($^2$H NMR) at 303~K~\cite{ulrich94}. 
    The signs are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
    and Gross et al.~\cite{gross97}.
    Note that to elucidate the relative change as a function of hydration level,
    the simulation results are shifted such that the (smaller) $S_\mathrm{CH}$
    matches (within $\pm$0.01) the experimental value at full hydration;
    the shift magnitudes for each of the force fields are listed ($S_\mathrm{CH}$+shift) in the y-label.
  }
\end{figure}

@@

We note that the modest ($\Delta S_\mathrm{CH}<0.02$ for g$_1$; $<0.04$ for g$_2$, g$_3$; see Fig.~\ref{ordPchol})
effects of cholesterol
on the glycerol backbone order parameters of POPC
measured by $^1$H-$^{13}$C NMR~\cite{ferreira13}
agree well with the results for phosphatidylethanolamine (PE) measured by $^{2}$H NMR~\cite{ghosh82}.
This further supports the ideas that the glycerol backbone structural behaviour is independent of the
headgroup composition~\cite{gally81} and that the headgroup structure is largely independent of the acyl chain region content unless
charges are present~\cite{scherer87}.
\begin{figure}[]
  \centering
  \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_lp.pdf}
%    \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_smooth.pdf}
  \caption{\label{ordPchol}
    The effect of cholesterol content on the glycerol backbone and choline order parameters in experiments~\cite{brown78,ferreira13} and simulations
    with the Berger-POPC-07/H\"oltje-CHOL-13, CHARMM36 and MacRog force fields. The signs in the experimental values are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
    and Gross et al.~\cite{gross97}.  Most order parameters from Berger-POPC-07/H\"oltje-CHOL-13 model for g$_1$ are beoynd the y-axis scale.
    In order to elucidate the relative change as a function of cholesterol content,
    the simulation results are shifted such that the (smaller) $S_\mathrm{CH}$
    matches (within $\pm$0.01) the experimental value without cholesterol;
    the shift magnitudes for each of the force fields are listed ($S_\mathrm{CH}$+shift) in the y-label.
}
\end{figure}

>>>>

\subsubsection{Phospholipid bilayer with low hydration level}
Fig.~\ref{ordPhydr} shows the published~\cite{dvinskikh05b,ulrich94,bechinger91} experimental order parameters
for the glycerol backbone and choline as a function of hydration level.
Despite slight differences in temperature and acyl chain composition,
the three independently reported data sets for the choline ($\beta$ and $\alpha$) segments agree well with each other:
Both order parameters increase with decreasing hydration level.
The glycerol backbone order parameters (g$_{3}$, g$_{2}$, g$_{1}$), in contrast,
have been observed~\cite{dvinskikh05b} to slightly decrease with dehydration.
Note that the original experiments~\cite{dvinskikh05b,ulrich94,bechinger91} measured only 
absolute values, but here
we included the signs measured in separate studies~\cite{hong95a,hong95b,gross97}. 
Consequently, the negative $\beta$ order parameter actually increases with dehydration 
as its absolute value decreases~\cite{dvinskikh05b,ulrich94,bechinger91}.
\begin{figure}[]
  \centering
  \includegraphics[width=7.3cm]{../DATAreportediINblog/dehydration_lp.pdf}
%    \includegraphics[width=7.3cm]{../DATAreportediINblog/dehydration_smooth.pdf}
    \caption{\label{ordPhydr}
    The effect of dehydration on glycerol and choline order parameters in experiments.
    The magnitudes of order parameters are measured for DMPC ($^1$H-$^{13}$C NMR) at 314~K~\cite{dvinskikh05b}, 
    for POPC ($^2$H NMR) at 296~K~\cite{bechinger91} and for DOPC ($^2$H NMR) at 303~K~\cite{ulrich94}. 
    The signs are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
    and Gross et al.~\cite{gross97}.
    Note that to elucidate the relative change as a function of hydration level, the simulation results were vertically shifted; 
    the shift magnitudes for each of the force fields are listed (SCH+shift) in the y-label.
  }
\end{figure}

@@

We note that the modest ($\Delta S_\mathrm{CH}<0.02$ for g$_1$; $<0.04$ for g$_2$, g$_3$; see Fig.~\ref{ordPchol})
effects of cholesterol
on the glycerol backbone order parameters of POPC
measured by $^1$H-$^{13}$C NMR~\cite{ferreira13}
agree well with the results for phosphatidylethanolamine (PE) measured by $^{2}$H NMR~\cite{ghosh82}.
This further supports the ideas that the glycerol backbone structural behaviour is independent of the
headgroup composition~\cite{gally81} and that the headgroup structure is largely independent of the acyl chain region content unless
charges are present~\cite{scherer87}.
\begin{figure}[]
  \centering
  \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_lp.pdf}
%    \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_smooth.pdf}
  \caption{\label{ordPchol}
    The effect of cholesterol content on the glycerol backbone and choline order parameters in experiments~\cite{brown78,ferreira13} and simulations
    with the Berger-POPC-07/H\"oltje-CHOL-13, CHARMM36 and MacRog force fields. The signs in the experimental values are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
    and Gross et al.~\cite{gross97}.  Most order parameters from Berger-POPC-07/H\"oltje-CHOL-13 model for g$_1$ are beoynd the y-axis scale.
    In order to elucidate the relative change as a function of cholesterol content, the simulation results were vertically shifted; 
    the shift magnitudes for each of the force fields are listed (SCH+shift) in the y-label.
}
\end{figure}
